---
title: "1_data_cleaning"
author: "sbsambado"
date: "5/23/2022"
output: html_document
---

# Multiparasite paper

## Title: Macro-parasites and micro-parasites co-exist in wildlife communities but are influencd by different community-level parameters

### Authors: Jordan Salomon, Samantha B Sambado, Arielle Crews, Suhkman Sidu, Eric Seredian, Adrienne Almarinez, Rachel Grgich, and Andrea Swei

#### Terminology

**Macro-parasites** : Ticks & helminths

**Micro-parasites**: Borrelia spp includes: BB = Borrelia burgdorferi sensu stricto, BBSL = Borrelia bissetti, BM = Borrelia miyamotoi, BARF = Bay Area Relapsing Fever)



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)

# needed packages
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(reshape2)
library(MASS)
library(lme4)
library(kableExtra)
library(psych)
library(effects)
library(jtools)
library(forcats)

```


**Data set 1**: `data` is the full data set from Swei database with 47 columns of data that is then subsetted to `data_subset` with just the variables needed for our analyses

```{r}

# read in data from Swei master data base (given to me by J. Salomon, need to check when this was accessed)
data <- read_csv("multiparasite.csv") # check out update Swei master list for NAs on tissue5S testing

# structure of data
dim(data) # 313 x 47
str(data) 

# clean up column names
names(data) <- tolower(names(data))

# check to make sure we're doing analyses on the sites we want
unique(data$site)

# select the columns wanted for these analyses
data_subset <- data[, -c(3,8,19,21, 23:38, 45:47)]   #c(1:7, 12:18, 21,22, 39, 40:44)]
str(data_subset)

# give column names that are easier to manipulate
names(data_subset)[2] <- "id"
names(data_subset)[5] <- "mass_log"
names(data_subset)[7] <- "taken_blood"
names(data_subset)[8] <- "tested_tissue5s"
names(data_subset)[9] <- "tested_tissue16s"
names(data_subset)[10] <- "path_presence" # bb or bm 
names(data_subset)[11] <- "hel_presence"
names(data_subset)[12] <- "hel_load"
names(data_subset)[13] <- "hel_rich"
names(data_subset)[14] <- "tic_presence"
names(data_subset)[15] <- "tic_load"
names(data_subset)[16] <- "tic_rich"
names(data_subset)[17] <- "path_rich" # bb or bm but double check
names(data_subset)[18] <- "macro_load"
names(data_subset)[19] <- "site_area_log"
names(data_subset)[20] <- "site_path_rich"
names(data_subset)[21] <- "site_vert_rich"
names(data_subset)[22] <- "site_rod_shan"
names(data_subset)[23] <- "site_pred_shan"
names(data_subset)[24] <- "site_qil_18"

# change to right data structures
data_subset$site <- as.factor(data_subset$site)
data_subset$genera <- as.factor(data_subset$genera)
data_subset$sex <- as.factor(data_subset$sex)
data_subset$age <- as.factor(data_subset$age)
data_subset$path_presence <- as.numeric(data_subset$path_presence)
data_subset$hel_presence <- as.numeric(data_subset$hel_presence)
data_subset$tic_presence <- as.numeric(data_subset$tic_presence)
data_subset$path_presence <- as.numeric(data_subset$path_presence)


# create log of counts
data_subset$hel_load_log <- log(data_subset$hel_load + 1)
data_subset$tic_load_log <- log(data_subset$tic_load + 1)

# check the transformation worked
str(data_subset) # 313 x 26


# write to a csv for future analyses
write.csv(data_subset, "data/data_subset.csv", row.names = FALSE)
```

**Data set 2**: `data_partial` is the subset of data that has just *N. fuscipes* woodrats (ie NEFU) and *Peromyscus spp.* mice (ie PERO). This dataset will be used for our analyses due to the consistent sampling of those genera. 

```{r}

# only going to select NEFU and PERO

data_partial <- data_subset[which(data_subset$genera == "NEFU" | data_subset$genera == "PERO"),]

# check that worked
unique(data_subset$genera)

str(data_partial) #304 x 26

# write to a csv for future analyses
write.csv(data_partial, "data/data_partial.csv", row.names = FALSE)
```

**Data set 3**: `prism_full` comes from a script I generated to scrape data from PRISM (NEED TO UPLOAD THIS) that will be merged into `data_complete_prism` that has all Swei data plus PRISM data

```{r}
# upload PRISM data
prism_full <- read.csv("data/PRISM_FULL_sites.csv")

# merge PRISM with Swei data

data_complete_prism <- merge(data_partial, prism_full, by = "site")

# write to a csv for future analyses
write.csv(data_complete_prism , "data/data_complete_prism.csv", row.names = FALSE)
```


**Data set 4**: `data_select` comes with the most recent pathogen data available on the Swei database that is then merged with `data_complete_prism`.

```{r}
master <- read.csv("2018_Master_Field_Data_pulled20220501.csv")

# select the stuff we want
master_subset <- master[, c(3,7, 8, 94)] 

#names(master_subset)[1] <- "site"
names(master_subset)[1] <- "date"
names(master_subset)[2] <- "id"
names(master_subset)[3] <- "species"
names(master_subset)[4] <- "overall_infection"

# checknames of infection
unique(master_subset$overall_infection)


# correct names
master_subset$overall_infection <- recode_factor(master_subset$overall_infection,
                                                 Neg = "Negative",
                                                 "Borrelia sp. (Lake Gaillard)" = "BARF",
                                                 "Borrelia sp. (novel relapsing fever)" = "BARF",
                                                 "MISSING" = "NA",
                                                 "Blood negative; ear punch needs to be extracted and tested" = "NA",
                                                 "0" = "NA",
                                                 "Bbsl (redo?)" = "BBSL",
                                                 "B. bissetti" = "BBSL",
                                                 "B. burgdorferi s.l. (not sequence confirmed, but strong PCR band)" = "BBSL",
                                                 "Neg? " = "Negative",
                                                 "<NA>" = "NA",
                                                 " " = "NA",
                                                 "missing" = "NA",
                                                 "B.burgdorferi, Borrelia sp. (Lake Gaillard)" = "CO:BBSS:BARF",
                                                 "B. miyamotoi, Borrelia sp. (novel relapsing fever)" = "CO:BM:BARF",
                                                 "B. burgdorferi, Borrelia sp. (novel relapsing fever)" = "CO:BBSS:BARF",
                                                 "B. burgdorferi" = "BB",
                                                 "B. miyamotoi" = "BM",
                                                 "B. burgdorferi " = "BB")

# check to make sure that worked
unique(master_subset$overall_infection)

# change nefu species
master_subset$species <- recode_factor(master_subset$species,
                                       "NEFU " = "NEFU")

# let's eliminate the rest
master_partial <- master_subset[which(master_subset$overall_infection == "BB" | 
                                        master_subset$overall_infection == "BM" |
                                        master_subset$overall_infection == "BBSL" |
                                        master_subset$overall_infection == "BARF" |
                                        master_subset$overall_infection == "CO:BBSS:BARF" |
                                        master_subset$overall_infection == "CO:BM:BARF" |
                                        master_subset$overall_infection == "Negative" |
                                        master_subset$overall_infection == "NA"),]

# check if that worked
str(master_partial) # 396 x 5

unique(master_partial$overall_infection)

# lets omit some columns to make it easier
unique(master_subset$species)

         
# now match it up with data frame by id
data_complete <- merge(data_partial, master_partial, by = "id") #260 x 29


# reorder the factor (for visualization purposes later)
data_complete$overall_infection <- fct_rev(data_complete$overall_infection)

# make the dataset without BARF
data_select <- data_complete[which(data_complete$overall_infection == "BB" | 
                                        data_complete$overall_infection == "BM" |
                                        data_complete$overall_infection == "BBSL" |
                                     data_complete$overall_infection == "Negative"),]

# check if that worked
unique(data_select$overall_infection)

# write to a csv for future analyses
write.csv(data_select, "data/data_select.csv", row.names = FALSE)

write.csv(data_complete, "data/data_complete.csv", row.names = FALSE)
```
































