---
title: "2_data_analysis"
author: "sbsambado"
date: "5/23/2022"
output: html_document
---

# Multiparasite paper

## Title: Macro-parasites and micro-parasites co-exist in wildlife communities but are influencd by different community-level parameters

### Authors: Jordan Salomon, Samantha B Sambado, Arielle Crews, Suhkman Sidu, Eric Seredian, Adrienne Almarinez, Rachel Grgich, and Andrea Swei

#### Terminology

**Macro-parasites** : Ticks & helminths

**Micro-parasites**: Borrelia spp includes: BB = Borrelia burgdorferi sensu stricto, BBSL = Borrelia bissetti, BM = Borrelia miyamotoi, BARF = Bay Area Relapsing Fever)

### Summary statistics


**This is for significant regression models only in Table X and X**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(message = FALSE)

# needed packages
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(reshape2)
library(MASS)
library(lme4)
library(kableExtra)
library(psych)
library(effects)
library(jtools)
library(forcats)
```


### Summary statistics

wilcoxon rank sum test
```{r}

## burden intensity for all rodents
wilcox.test(data_partial$tic_load ~ data_partial$genera) # W = 12806, p-value = 2.033e-05

wilcox.test(data_partial$hel_load ~ data_partial$genera) # W = 1984.5, p-value = 1.598e-06


## burden intensity for infected and unifected NEFU

nefu <- data_partial %>%
  filter(genera == "NEFU")


wilcox.test(nefu$tic_load ~ nefu$path_presence) # W = 596, p-value = 0.2878

wilcox.test(nefu$hel_load ~ nefu$path_presence) # W = 249, p-value = 0.9905


wilcox.test(tic_load ~ path_presence , data = data_partial) # W = 5263, p-value = 0.03019

wilcox.test(hel_load ~ path_presence, data = data_partial) # W = 1515, p-value = 0.0406


# calculate mean tick & helminth load for all rodents

mean(data_partial$tic_load[data_partial$path_presence == 0], na.rm = TRUE) # 1.636735
mean(data_partial$tic_load[data_partial$path_presence == 1], na.rm = TRUE) # 2.403846
## for all rodents, the mean tick load was higher in infected (2.4) than non-infected (1.6)

mean(data_partial$hel_load[data_partial$path_presence == 0], na.rm = TRUE) # 4.648148
mean(data_partial$hel_load[data_partial$path_presence == 1], na.rm = TRUE) # 1.489362
## for all rodents, the mean helminth load was higher in non-infected (4.6) than infected

# calculate mean tick & helminth load for NEFU
mean(nefu$tic_load[nefu$path_presence == 0], na.rm = TRUE) # 3.783784
mean(nefu$tic_load[nefu$path_presence == 1], na.rm = TRUE) # 4.894737
## for NEFUs, the mean tick load was higher in infected (4.9) than non-infected (3.8)

mean(nefu$hel_load[nefu$path_presence == 0], na.rm = TRUE) # 7.774194
mean(nefu$hel_load[nefu$path_presence == 1], na.rm = TRUE) # 4.1875
## for NEFUs, the mean helminth load was higher in non-infected (7.8) than infected (4.2)

```



### Part 1: Characterize the distribution of macro- and micro- parasites across individual hosts. 


###### Rodent tick burden (negative binomial model) with site as random effect
```{r}
## build models
mod_1_tic <- glmer.nb(tic_load  ~ genera + sex + age + (1|site),
                    data = data_partial)


## model summary
summary(mod_1_tic)


## model diagnostics
# determine if model is significant without genera
mod_1_tic2 <- update(mod_1_tic,.~. - genera)
anova(mod_1_tic, mod_tic2) # genera is a significant predictor p = 6.799e-13 ***


### model visualization
plot(predictorEffects(mod_1_tic))


```

###### Rodent helminth burden (negative binomial model) with site as random effect
```{r}
## build model
mod_1_hel <- glmer.nb(hel_load ~ genera + sex + age + (1|site),
                    data = data_partial) 

### model summary
summary(mod_1_hel)


## model diagnostics
# determine if model is significant without genera
mod_1_hel2 <- update(mod_1_hel,.~. - genera)
anova(mod_hel, mod_hel2) # genera is a significant predictor  p = 9.699e-08 ***


### model visualization
plot(predictorEffects(mod_1_hel))


### compare model coefficients
mod_1_coefficients_plot <- plot_summs(mod_1_hel, mod_1_tic,
           model.names = c("Helminth" , "Tick"),
           #plot.distributions = TRUE,
           #scale = TRUE,
           inner_ci_level = .8,
           coefs = c("Genera - Peromyscus" = "generaPERO",
                     "Sex - Male" = "sexM",
                     "Age - Juvenile" = "ageJ")
           )

```

###### Rodent B. burgdorferi sensu lato infection (binomial model) 

```{r}
## select for that data
data_select_bbsl <- data_select %>%
  filter(overall_infection == "BB" | overall_infection == "BBSL" |overall_infection == "Negative")


## build model 
mod_1_logit_bbsl <- glm(path_presence ~ genera + sex + age, 
                         family = "binomial", data = data_select_bbsl)


## model summary
summary(mod_1_logit_bbsl)


### model visualization
plot(predictorEffects(mod_1_logit_bbsl))


```

### Part 2: Understand how macro-parasites are influenced by host community dynamics

###### Tick burden (negative binomial model) with genera as random effect
```{r}
## build model
mod_2_tic_community <- glmer.nb(tic_load ~ site_rod_shan + site_pred_shan + (1|genera),
                data = data_partial)


## model summary
summary(mod_2_tic_community)


## visualize model
plot_summs(mod_2_tic_community,
           inner_ci_level = .8,
           coefs = c("Site Rodent Shannon Diversity" = "site_rod_shan",
                     "Site Predator Shannon Diversity" = "site_pred_shan")
)


```

###### Tick burden (negative binomial model) with genera as random effect for mean temperature (C)

```{r}
## select data
climate_macro <- data_complete_prism %>%
  group_by(site,data_type) %>%
  dplyr::select(site, genera, tic_load, hel_load, data)


## filter mean temperature
tmean_macro <- climate_macro %>%
  filter(data_type == "tmean")

## build model
mod_2_tic_tmean <- glmer.nb(tic_load ~ data + (1|genera), data = tmean_macro)


# model summary 
summary(mod_2_tic_tmean)


```

###### Tick burden (negative binomial model) with genera as random effect for max vapor pressure deficit

```{r}
## select data
climate_macro <- data_complete_prism %>%
  group_by(site,data_type) %>%
  dplyr::select(site, genera, tic_load, hel_load, data)


## filter vpdmax
vpdmax_macro <- climate_macro %>%
  filter(data_type == "vpdmax")


## build model
mod_2_tic_vpdmax <- glmer.nb(tic_load ~ data + (1|genera), data = vpdmax_macro) # significant (estimate = 0.11075   sd =   0.04899  z-score =   2.261   p-value = 0.0238)

## model summary
summary(mod_2_tic_vpdmax)


```

###### Pathogen presence (binomial model) for N. fuscipes only

```{r}
## select data
 nefu_all <- data_select %>%
  filter(genera == "NEFU") 


## build model
mod_2_logit_nefu_all <- glm(path_presence ~ site_rod_shan + site_pred_shan,
                data = nefu_all, 
                family = "binomial")


## model summary
summary(mod_2_logit_nefu_all)


```

### Part 3: Analyze how macro-parasites might influence micro-parasite infection within individual hosts

There were no significant models for these analyses on this data set.